---
title: "Clustering with code"
output: html_notebook
---

# 1. Overview
This notebook is to save all information how to do clustering with code using k-means algorithm. 

# 2. Data
Same data as used to do clustering by hand - Lambeth Borough demographic data, including age and income

# 3. Preparation
Install and load libraries needed to finish it (as per Kyle's webinar).
Tidyverse


```{r}
library(tidyverse)
library(readr)
library(ggplot2)
```

# 4. Load data
Loading data for analysis and store them under "data" variable

```{r}
data <- read.csv("Clusteringdata.csv")
```

# 5. Explore and transform the data
Check if data loaded correctly using head function
Rename columns
Change data types to the correct one - Ward = string, Age - double, Income - integer

```{r}
names(data) <- c("Ward", "Age", "Income")
data$Ward <- as.character(data$Ward)
head(data)
```


If the numeric data has got any other symbol (like currency), "parse" function can be used to get rid of all-numeric symbols

```{r}
data$Income <- parse_number(data$Income)
```

#6. Visualise the data
This is to plot a graph that will show the data using qplot function. Other functions can be used as well like qqplot

```{r}
qplot(x = data$Age, y = data$Income, label=data$Ward, geom = "text")
ggplot(data)+
  geom_point(data = data)+
  aes(data$Age, data$Income)
```


# 7. Clustering
This is when k-means is used. 


First we need to determine the number of clusters - arbitrarily it's 3.

```{r}
NoClusters <- 3
```





Then we need to run k-means function to find centroids in the dataset and assign it to the new value. Each ward will be assigned to one centroid (group)
```{r}
kMeansResult <- kmeans(data[c("Age", "Income")], centers = NoClusters)
```





Then we will create a Cluster column in the dataset and copy values found by k-means. 
```{r}
data$Clusters <- c(kMeansResult$cluster)
head(data)
```





Then we need to convert the Clusters column to a factor
```{r}
data$Clusters <- as.factor(data$Clusters)
```


# 8. Visualise a cluster
To visualise a cluster, the same qplot function will be used with an additional argument - each color will represent a group as per the result of k-means

```{r}
qplot(x=data$Age, y=data$Income, label=data$Ward, geom = "text", color=data$Clusters)
```


# 9. (Optional) Fine tuning cluster number
Number of clusters in k-means depends on the total of within-cluster sum of squares (WSS) . The optimal number of clusters can be found by plotting a function that will show the the points for 1 to n clusters with a value of WSS. Thanks to this you can find an elbow (bend) which indicates the optimal k number.

```{r}
kClustersMax <- 10

wss <- sapply(1:kClustersMax,
        function(k){kmeans(data[c("Age", "Income")], k)$tot.withinss})

plot(1:kClustersMax, wss,
       type="b", pch = 19, frame = FALSE,
       xlab="Number of clusters K",
       ylab="Total within-clusters sum of squares")
```

In our case the optimal value is 2 or 3 - depending on the interpretation. 

